---
title: "In-class Exercise 6"
description: |
  A short description of the post.
author:
  - name: Ngah Xin Yan
    url: https://github.com/nxinyan/
date:  09-20-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.retina = 3)
```

## Overview 

In this hands-on exercise, you will learn how to compute spatial weights using R. By the end to this hands-on exercise, you will be able to:

- import geospatial data using appropriate function(s) of **sf** package

- import csv file using appropriate function of **readr** package

- perform relational join using appropriate join function of **dplyr** package

- compute spatial weights using appropriate functions of **spdep** package

- calculate spatially lagged variables using appropriate functions of **spdep** package.

## Data Used

The following datasets were used:

- Hunan county boundary layer. This is a geospatial data set in ESRI shapefile format.

- Hunan_2012.csv: This csv file contains selected Hunan’s local development indicators in 2012.

## Installing and Loading the R packages

```{r}
packages = c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Data Import and Preparation

The code chunk below uses *st_read()* of **sf** package to import Hunan shapefile into R. The imported shapefile will be **simple features** Object of **sf**.

### Import shapefile into r environment

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### Import csv file into r environment

Next, we will import Hunan_2012.csv into R by using read_csv() of readr package. The output is R dataframe class.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

### Performing relational join

The code chunk below will be used to update the attribute table of *hunan*’s SpatialPolygonsDataFrame with the attribute fields of *hunan2012* dataframe. This is performed by using *left_join()* of **dplyr** package.

```{r}
hunan <- left_join(hunan,hunan2012)
```

### Visualising Regional Development Indicator

Now, we are going to prepare a choropleth map showing the distribution of GDPPC 2012 by using *qtm()* of **tmap** package.

** In-class Exercise**

```{r fig.width=12, fig.height=8}
basemap <- tm_shape(hunan) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan, "GDPPC") #quick tmap method
tmap_arrange(basemap, gdppc, asp=1, ncol=2)

```

- Display map with location name and all county names are inserted in on the base map, so can reference a particular county from right map with left map
- Show use of tmap to plot a basemap, show how to insert text onto polygon

# Computing Contiguity Spatial Weights

In this section, you will learn how to use *poly2nb()* of **spdep** package to compute contiguity weight matrices for the study area. This function builds a neighbours list based on regions with contiguous boundaries. If you look at the documentation you will see that you can pass a “queen” argument that takes TRUE or FALSE as options. If you do not specify this argument the default is set to TRUE, that is, if you don’t specify queen = FALSE this function will return a list of first order neighbours using the Queen criteria.

## Computing (QUEEN) contiguity based neighbours

```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```

The summary report above shows that there are 88 area units in Hunan. The most connected area unit has 11 neighbours. There are two area units with only one neighbours.

For each polygon in our polygon object, wm_q lists all neighboring polygons. For example, to see the neighbors for the first polygon in the object, type:

```{r}
wm_q[[1]]
```

Polygon 1 has 5 neighbors. The numbers represent the polygon IDs as stored in hunan SpatialPolygonsDataFrame class.

We can retrive the county name of Polygon ID=1 by using the code chunk below:

```{r}
hunan$County[1]
```

The output reveals that Polygon ID=1 is Anxiang county.

To reveal the county names of the five neighboring polygons, the code chunk will be used:

```{r}
hunan$NAME_3[c(2,3,4,57,85)]
```

** In-class Exercise**
We can retrieve the GDPPC of these five countries by using the code chunk below.
- purpose: find out the neighbour's gdp per capital

```{r}
nb1 <- wm_q[[1]]
nb1 <- hunan$GDPPC[nb1]
nb1
```
The printed output above shows that the GDPPC of the five nearest neighbours based on Queen's method are 20981, 34592, 24473, 21311 and 22879 respectively.

You can display the complete weight matrix by using str().

```{r}
str(wm_q)
```

The summary report above shows that there are 88 area units in Hunan. The most connect area unit has 10 neighbours. There are two area units with only one heighbours.

## Creating (ROOK) contiguity based neighbours

```{r}
wm_r <- poly2nb(hunan, queen=FALSE)
summary(wm_r)
```

## Visualising contiguity weights

To get our longitude values we map the st_centroid function over the geometry column of us.bound and access the longitude value through double bracket notation [[]] and 1. This allows us to get only the longitude, which is the first value in each centroid.

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
```

- st_centroid(): calculate centroid of polygons, center point of each geographical area mathematically, so we can use them

- first one is long, second one is lat, coordinate system is lat,long, but computer store long (x-axis) then lat(y-axis), start with x-axis then y-axis

- once retrieve their centroid, combine them to be coordinate and plot it

We do the same for latitude with one key difference. We access the second value per each centroid with [[2]].

```{r}
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
```

Now that we have latitude and longitude, we use cbind to put longitude and latitude into the same object.

```{r}
coords <- cbind(longitude, latitude)
```

We check the first few observations to see if things are formatted correctly.

```{r}
head(coords)
```

### Plotting Queen contiguity based neighbours maps

```{r fig.width=8, fig.height=6}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```


### Plotting Rook contiguity based neighbours maps

```{r fig.width=8, fig.height=6}
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

```{r fig.width=8, fig.height=6}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red", main="Queen Contiguity")
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red", main="Rook Contiguity")
```

## Computing distance based neighbours

- conversion during calculation to km, cannot calculate distance with decimal degree, as long as is lat long, it will return as km

### Determine the cut-off distance

Firstly, we need to determine the upper limit for distance band by using the steps below:

- Return a matrix with the indices of points belonging to the set of the k nearest neighbours of each other by using knearneigh() of spdep.

- Convert the knn object returned by knearneigh() into a neighbours list of class nb with a list of integer vectors containing neighbour region number ids by using knn2nb().

- Return the length of neighbour relationship edges by using nbdists() of spdep. The function returns in the units of the coordinates if the coordinates are projected, in km otherwise.

- Remove the list structure of the returned object by using unlist().

```{r}
#coords <- coordinates(hunan)
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

The summary report shows that the largest first nearest neighbour distance is 61.79 km, so using this as the upper threshold gives certainty that all units will have at least one neighbour.
   z    
### Computing fixed distance weight matrix

Compute the distance weight matrix by using dnearneigh() 

```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

Next, we will use str() to display the content of wm_d62 weight matrix.

```{r}
str(wm_d62)
```

Another way to display the structure of the weight matrix is to combine table() and card() of spdep.

```{r}
table(hunan$County, card(wm_d62))
```

```{r}
n_comp <- n.comp.nb(wm_d62)
n_comp$nc
```

```{r}
table(n_comp$comp.id)
```

### Plotting fixed distance weight matrix

Plot distance weight matrix

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE)
plot(k1, coords, add=TRUE, col="red", length=0.08)
```

The red lines show the links of 1st nearest neighbours and the black lines show the links of neighbours within the cut-off distance of 62km.

Plot both of them next to each other

```{r fig.width=8, fig.height=6}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(k1, coords, add=TRUE, col="red", length=0.08, main="1st nearest neighbours")
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE, pch = 19, cex = 0.6, main="Distance link")
```

### Computing adaptive distance weight matrix

One of the characteristics of fixed distance weight matrix is that more densely settled areas (usually the urban areas) tend to have more neighbours and the less densely settled areas (usually the rural counties) tend to have lesser neighbours. Having many neighbours smoothes the neighbour relationship across more neighbours.

It is possible to control the numbers of neighbours directly using k-nearest neighbours, either accepting asymmetric neighbours or imposing symmetry as shown in the code chunk below.

```{r}
knn6 <- knn2nb(knearneigh(coords, k=6))
knn6
```

```{r}
str(knn6)
```

Notice that each county has six neighbours, no less no more!

### Plotting distance based neighbours

Plotting weight matrix

```{r fig.width=8, fig.height=6}
plot(hunan$geometry, border="lightgrey")
plot(knn6, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## Weights based on IDW

In this section, you will learn how to derive a spatial weight matrix based on Inversed Distance method.

First, we will compute the distances between areas by using nbdists() of spdep.

```{r}
dist <- nbdists(wm_q, coords, longlat = TRUE)
ids <- lapply(dist, function(x) 1/(x))
ids
```

## Row-standardised weights matrix

```{r}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```

The zero.policy=TRUE option allows for lists of non-neighbors. This should be used with caution since the user may not be aware of missing neighbors in their dataset however, a zero.policy of FALSE would return an error.

To see the weight of the first polygon’s four neighbors type:

```{r}
rswm_q$weights[10]
```

Each neighbor is assigned a 0.2 of the total weight. This means that when R computes the average neighboring income values, each neighbor’s income will be multiplied by 0.2 before being tallied.

Derive a row standardised distance weight matrix 

```{r}
rswm_ids <- nb2listw(wm_q, glist=ids, style="B", zero.policy=TRUE)
rswm_ids
```

```{r}
rswm_ids$weights[1]
```

```{r}
summary(unlist(rswm_ids$weights))
```
## Application of Spatial Weight Matrix

In this section, you will learn how to create four different spatial lagged variables, they are:

- spatial lag with row-standardized weight

- spatial lag as a sum of neighbouring values

- spatial window average, and spatial window sum

### Spatial lag with row-standardized weights

Finally, we’ll compute the average neighbor GDPPC value for each polygon. These values are often referred to as spatially lagged values.

```{r}
GDPPC.lag <- lag.listw(rswm_q, hunan$GDPPC)
GDPPC.lag
```

Recalled in the previous section, we retrieved the GDPPC of these five countries by using the code chunk below.
```{r}
nb1 <- wm_q[[1]]
nb1 <- hunan$GDPPC[nb1]
nb1
```

We can append the spatially lag GDPPC values onto hunan SpatialPolygonDataFrame by using the code chunk below.

```{r}
# run separately to find out what each line does
lag.list <- list(hunan$NAME_3, lag.listw(rswm_q, hunan$GDPPC))
```

```{r}
lag.res <- as.data.frame(lag.list)
```

```{r}
colnames(lag.res) <- c("NAME_3", "lag GDPPC")
```

```{r}
hunan <- left_join(hunan,lag.res)
```

```{r}
head(hunan)
```

Next, we will plot both the GDPPC and spatial lag GDPPC for comparison using the code chunk below.

```{r fig.width=8, fig.height=6}
gdppc <- qtm(hunan, "GDPPC")
lag_gdppc <- qtm(hunan, "lag GDPPC")
tmap_arrange(gdppc, lag_gdppc, asp=1, ncol=2)
```
## Spatial window sum

The spatial window sum uses and includes the diagonal element. To begin, we will assign knn6 to a new variable because we will directly alter its structure to add the diagonal elements

```{r}
knn6a <- knn6
```

To add the diagonal element to the neighbour list, we just need to use include.self() from spdep

```{r}
include.self(knn6a)
```

Next, we will assign binary weights to the neighbour structure that includes the diagonal element.

```{r}
binary.knn6 <- lapply(knn6a, function(x) 0*x+1)
binary.knn6[1]
```

Use nb2listw() and glist() to explicitly assign weight values

```{r}
wm_knn6 <- nb2listw(knn6a, glist = binary.knn6, style = "B")
```

Compute the lag variable with lag.listw()

```{r}
lag_knn6 <- lag.listw(wm_knn6, hunan$GDPPC)
```

Convert the lag variable listw object into a data.frame by using as.data.frame()

```{r}
lag.list.knn6 <- list(hunan$NAME_3, lag.listw(wm_knn6, hunan$GDPPC))
lag_knn6.res <- as.data.frame(lag.list.knn6)
colnames(lag_knn6.res) <- c("NAME_3", "lag_sum GDPPC")
```

Append lag_sum GDPPC values onto hunan sf data.frame by using left_join() of dplyr package

```{r}
hunan <- left_join(hunan, lag_knn6.res)
```

Lastly, qtm() of tmap package is used to plot the GDPPC and lag_sum GDPPC map next to each other for quick comparison.

```{r fig.width=12, fig.height=8}
gdppc <- qtm(hunan, "GDPPC")
lag_sum_gdppc <- qtm(hunan, "lag_sum GDPPC")
tmap_arrange(gdppc, lag_sum_gdppc, asp=1, ncol=2)
```
