---
title: "In-Class Exercise 5"
description: |
  A short description of the post.
author:
  - name: Ngah Xin Yan
    url: https://github.com/nxinyan/
date: 09-13-2021
output:
  distill::distill_article:
    self_contained: false
---

Global Setting

```{r setup, include=FALSE, eval=TRUE, echo=TRUE, message=FALSE, error=FALSE, fig.retina=3}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing and Loading the R package

- maptools is for creating spatial* objects

- sf is for importing geospatial data

- raster is for creating raster object

- spatstat is for performing SPPA

- tmap is for plotting thematic maps

- tidyverse is for data handling

```{r}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap','tidyverse','plotly','ggthemes' )
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
```

## Importing the Geospatial Data

Importing shapefile using *st_read()* of **sf** package. The output object is in tittle sf object class.

```{r}
mpsz_sf <- st_read(dsn = "data/shapefile", 
                layer = "MP14_SUBZONE_WEB_PL")
```
Projection is in SVY21

## Importing aspatial data from rds folder

**read_rds()** of **readr** package is used instead of **readRDS()** of base R is used. This is because output of **read_rds()** is in tibble object

```{r}
childcare <- read_rds("data/rds/childcare.rds")
CHAS <- read_rds("data/rds/CHAS.rds")
```

Note that there are some data issue in childcare data frame because `Lat` and `Lng` should be in numeric data type. The coordinate fields are in decimal degrees. Hence, wgs referencing system is assumed.

## Converting the aspatial data frame into sf objects

```{r}
CHAS_sf <- st_as_sf(CHAS,
                         coords = c("X_COORDINATE",
                                    "Y_COORDINATE"),
                         crs=3414)
```

Note: *st_as_sf* accept coordinates in character data type.

```{r}
childcare_sf <- st_as_sf(childcare,
                         coords = c("Lng",
                                    "Lat"),
                         crs=4326) %>%
  st_transform(crs = 3414)
```

## Plotting for reviewing childcare and CHAS

```{r}
tmap_mode('view')
tm_shape(childcare_sf) +
  tm_dots(alpha = 0.4,
          col = "blue",
          size = 0.05) +
tm_shape(CHAS_sf) +
  tm_dots(alpha = 0.4,
          col = "red",
          size = 0.05)
```

## Geospatial Data Wrangling

### Converting frpm sf to Spatial* Classes

as_Spatial() of sf package.

```{r}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
```

### Convert Spatial* dataframe into Spatial* objects

as.SpatialPoint() of as.SpatialPolygon() of maptools package

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
mpsz_sp <- as(mpsz, "SpatialPolygons")
```

### Converting from Spatial* objecys into ppp objects

Using as.ppp() of maptools package, 

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
CHAS_ppp <- as(CHAS_sp, "ppp")
```

### Removing duplicate points using jitter

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp,
                             retry = TRUE,
                             nsim = 1,
                             drop = TRUE)
any(duplicated(childcare_ppp_jit))
```

```{r}
CHAS_ppp_jit <- rjitter(CHAS_ppp,
                             retry = TRUE,
                             nsim = 1,
                             drop = TRUE)
any(duplicated(CHAS_ppp_jit))
```
### Extracting Punggol Planning Area

```{r}
pg <- mpsz[mpsz@data$PLN_AREA_N=="PUNGGOL",]
```
### Converting SpatialPolygonsDataFrame into SpatialPolygons object

```{r}
pg_sp <- as(pg, "SpatialPolygons")
```

### Converting SpatialPolygons into owin object

```{r}
pg_owin <- as(pg_sp, "owin")
```

### Extracting spatial points within owin

```{r}
childcare_pg <- childcare_ppp_jit[pg_owin]
CHAS_pg <- CHAS_ppp_jit[pg_owin]
```

```{r}
plot(childcare_pg)
```

## L-function

```{r}
L_childcare <- envelope(childcare_pg,
                        Lest,
                        nsim=99,
                        rank=1,
                        global = TRUE)
```

